---
title: "kz2383_hw6"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


problem 2

data cleaning steps:create variable city_state, resolution ; eliminate cities with no race reported and wrong entry; limit race to black and white for analysis
```{r}

homicide = 
  read_csv("homicide.csv") %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    !(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) %>% 
  mutate(victim_race = fct_relevel(victim_race, "White")) %>% 
  select(city_state, resolution, victim_age, victim_sex, victim_race)


```


After data cleaning, there are  `r homicide%>% nrow()` cases recorded over the past decades. There are `r homicide%>% ncol()` variables included in the data. Possible NAs are dropped for better analysis. 


Compute 95% confidence interval for odds ratio of solving homocides comparing male and female in  baltimore and the each city in the complete dataset

```{r}

baltimore_glm = 
  filter(homicide, city_state == "Baltimore, MD") %>% 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = .)

baltimore_glm %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) %>% 
  filter(term == "victim_sexMale") %>% 
  select(OR, OR_CI_lower, OR_CI_upper)


results = 
  homicide %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = map(data, ~glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = .x)),
    tidy_models = map(models, broom::tidy)) %>% 
  select(-models, -data) %>% 
  unnest(tidy_models) %>% 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

results %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


The confidence interval for odds ratio of homicide case in baltimore is (0.325,0.558) with an estimated odds ratio of 0.426. 

Among the rest cities in the complete dataset. New York NY has the lowest odds ratio for comparing homicides solved between male and female victims with the upper confidence interval lower than 1 ,suggesting small difference between number of homicides solved for males and females. Albuquerque,NM has the highest odds ratio for comparing homicides solved comparing male victims and female victims with upper confidence interval almost reaching 3.8, suggesting huge difference between number of homicides solved for males and females. 







problem 3


data cleaning step：babysex, father race and mother race are recoded and malform is converted to logical value. 
```{r}

bwt = read_csv("birthweight.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as.factor(frace),
    frace = fct_recode(frace, "white" = "1", "black" = "2", "asian" = "3", 
                       "puerto rican" = "4", "other" = "8"),
    malform = as.logical(malform),
    mrace = as.factor(mrace),
    mrace = fct_recode(mrace, "white" = "1", "black" = "2", "asian" = "3", 
                       "puerto rican" = "4"))

bwt
```
Variable `babysex`,`blength`,`bhead`and `parity` are included in the linear model. Based on interpretation of the model, there is a negative intercept of -6125.3. All four variables have a positive estimated value and are significant with p-value much lower than 0.05. The plot displaying residuals Vs fitted values suggests a rather good fit for the model except for some data points distribute randomly away from the center. Specifically, the observed residual seems to be rather large for fitted value of 500 ,750 and around 1500.



```{r}
model1 = lm(bwt ~ babysex + blength + bhead + parity , data = bwt)

model1 %>% 
  broom::tidy()

bwt %>%
  add_residuals(model1) %>%
  add_predictions(model1) %>%
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  labs(
    x = "fitted values",
    y = "residuals",
    title = " residuals Vs fitted values plot"
  )
```
```{r}

library(purrr)

cv_df = 
  crossv_mc(bwt, 100) %>% 
  mutate(
    model2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model3 = map(train, ~lm(bwt ~ babysex + blength + bhead + parity, data = .x)),
    model4 =  map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex+ babysex * blength+ bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_mod2 = map2_dbl(model2, test, rmse),
    rmse_mod3 = map2_dbl(model3, test, rmse),
    rmse_mod4 = map2_dbl(model4, test, rmse)) 

cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```
Comparing root mean square error for each constructed model, model 3 includes `babysex`,`blength`,`bhead`and `parity` has the lowest value, suggesting the best fitness. The first model only includes`blenght`and `gaweeks` has the largest rmse as expected due to limited number of estimators taken into account. The last model includes `bhead`,`blength`, `babysex` and all possible variable interactions has slightly higher rmse than the second model. This could potentially be explained by existing interactions between variables. In general. model3 turned out to be the optimal one after comparing rmse values, suggesting the best fit among the three. 
